---
title: "data_exploration"
output: html_document
date: "2023-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries

```{r}
library(tidyverse)
```

## Loading CSV Data

Inside the utility directory holding the data, there are a number of csv files that describe the dataset. We will practice loading these files, and demonstrate some very simple plotting, to understand what data is available to us throughout the tutorial.

```{r}
tcell_summary <- read_csv("../../utility/summaryInfo/TcellSummaryTable.csv")
tcell_summary
```

At this point, you should a *tcell_summary* tibble object in your global environment.

## Long and Wide Format

The current data frame is in **wide** format. To convert to tables that are in **long** format, we can use the **pivot_longer** function from tidyverse.

```{r}
tissue <- tcell_summary["...1"]
tcell_summary <- select(tcell_summary, -c("...1"))
source = colnames(tcell_summary)
tcell_summary["Tissue"] <- tissue
tcell_table <- pivot_longer(tcell_summary, names_to = "Source", values_to = "Count", cols=source)
```

After converting to long format.

```{r tcell_table}
tcell_table
```

## Bar plots

Let's plot the number of T cells across tissue types in the cohort by source.

```{r bar, echo=FALSE}
ggplot(tcell_table, aes(x=Tissue, y=Count, fill=Source)) + geom_bar(stat='identity', position='dodge') + theme(aspect.ratio=1/3, axis.text.x = element_text(angle=45,vjust=1,hjust=1))
```


## Heatmap

Let's plot the same thing as a heatmap.

```{r heatmap, echo=FALSE}
ggplot(tcell_table, aes(Source, Tissue, fill= Count)) + 
    geom_tile() +
    scale_fill_gradient(low="white", high="orange")
```
